---
title: "Homework2"
author: "Ethan Marburger"
format: html
editor: visual
---

## Homework 2

### Step 3

#### Task 1: Basic Vector Practice

1A) Creating a vector from pre-treatment blood pressure values.

```{r}
pretreatment <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
pretreatment
```
1B) Creating a vector from post-treatment blood pressure values.

```{r}
posttreatment <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
posttreatment
```
2) Assigning names to vector elements using the name() and paste() functions

```{r}
names(pretreatment) <- paste("Subject", 1:20, sep = "_")
names(posttreatment) <- paste("Subject", 1:20, sep = "_")
```

3) Calculate the change in blood pressure for each patient by subtracting post-treatment measurements
from pre-treatment measurements.

```{r}
bloodpressure_dif <- pretreatment - posttreatment
bloodpressure_dif
```

4) Calculate the average decrease in blood pressure across all patients.

```{r}
mean_bp_dif <- mean(bloodpressure_dif)
mean_bp_dif
```
5) Determine which patients experienced a decrease in blood pressure after treatment (a positive change).

```{r}
positive_change_indices <- which(bloodpressure_dif > 0)
positive_change_indices
```

6) Subsetting the vector of differences to only return those that have a positive change.

```{r}
boolean_postiive_bp <- bloodpressure_dif[bloodpressure_dif > 0]
boolean_postiive_bp
```

7) Calculating the average decrease in blood pressure for those where the blood pressure decreased (positive
change)

```{r}
mean_positive_bp_dif <- mean(boolean_postiive_bp)
mean_positive_bp_dif
```


#### Task 2: Basic Data Frame Practice

1) Creating a data frame using previous data

```{r}
patient <- paste("Subject", 1:20, sep = "_")
df <- data.frame(patient, pre_bp = pretreatment, post_bp = posttreatment, diff_bp = bloodpressure_dif)
df
```

2) Returning only rows of the data frame where the diff_bp column is negative

```{r}
df_positive_dif <- df[df$diff_bp > 0, ]
df_positive_dif
```

3) Adding a new column to my data frame orresponding to TRUE if the post_bp is less than 120 and FALSE
if it is not

```{r}
df <- df[, -5] #ignore, deleting extra column
df$post_bp_120 <- df$post_bp < 120
df
```

4) Printing data frame nicely

```{r}
knitr::kable(df)
```


#### Task 3: List Practice

1) Creating a new data frame with this data similar to task 2.

```{r}
#Combing values
pre_bp <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_bp <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
#Difference between pre_trt and post_trt values
bp_dif <- pre_bp - post_bp
#Defining patient column
patient <- paste("Subject", 1:20, sep = "_")
#Creating a data frame
placebo_df <- data.frame(patient, pre_bp, post_bp, bp_dif)
#Adding new column based on if the post_trt is above 120 or not
placebo_df$post_bp_120 <- placebo_df$post_bp < 120
placebo_df
```

2) Creating and storing both data frames in a list

```{r}
bp_list <- list(treatment = df, placebo = placebo_df)
bp_list
```

3) Accessing the first list element using three different types of syntax

```{r}
#Syntax 1
bp_list[1][[1]]
```

```{r}
#Syntax 2
bp_list[[1]]
```

```{r}
#Syntax 3
bp_list[1]
```

4) In one line of code, accessing the pre_bp column of the placebo data frame.

```{r}
bp_list[[2]][["pre_bp"]]
```

