---
title: "Homework2"
author: "Ethan Marburger"
format: html
editor: visual
---

## Homework 2

### Step 3

#### Task 2: Basic Vector Practice

1A) Creating a vector from pre-treatment blood pressure values.

```{r}
pretreatment <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
pretreatment
```
1B) Creating a vector from post-treatment blood pressure values.

```{r}
posttreatment <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
posttreatment
```
2) Assigning names to vector elements using the name() and paste() functions

```{r}
names(pretreatment) <- paste("Subject", 1:20, sep = "_")
names(posttreatment) <- paste("Subject", 1:20, sep = "_")
```

3) Calculate the change in blood pressure for each patient by subtracting post-treatment measurements
from pre-treatment measurements.

```{r}
bloodpressure_dif <- pretreatment - posttreatment
bloodpressure_dif
```

4) Calculate the average decrease in blood pressure across all patients.

```{r}
mean_bp_dif <- mean(bloodpressure_dif)
mean_bp_dif
```
5) Determine which patients experienced a decrease in blood pressure after treatment (a positive change).

```{r}
positive_change_indices <- which(bloodpressure_dif > 0)
positive_change_indices
```

6) Subsetting the vector of differences to only return those that have a positive change.

```{r}
boolean_postiive_bp <- bloodpressure_dif[bloodpressure_dif > 0]
boolean_postiive_bp
```

7) Calculating the average decrease in blood pressure for those where the blood pressure decreased (positive
change)

```{r}
mean_positive_bp_dif <- mean(boolean_postiive_bp)
mean_positive_bp_dif
```